---
title: "FARS_Data"
author: "Tedros gebregergs"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FARS_Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = TRUE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Package Description
`farsdata` package loads data from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System(FARS).`FARS` data is  a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.  https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS).

# install.packages("devtools")
# devtools::install_github("hahustat/farsdata")

```{r setup}
library(farsdata)
library(dplyr)
library(maps)
```

# Loading/reading data

```{r}
list.files(system.file("extdata", package = "farsdata"))
```

## `make_filename` function
A function with input of year from 2013-2015 and returns a file and a path to the data files in the source package.
```{r}
year_input<-2013
make_filename(year_input)
```
## `fars_read_year` function
A function which takes an input of years and uses `make_file` function to create a file, and then reads for each year using a function `fars_read`. Result of this function will be used as input to function `fars_summarize_years`.
```{r}
library(dplyr)
fars_read_years(2015)
```
## `fars_summarize_years` function
Taking organized data by function `fars_read_year`, summarizes yearly accidents data by month.

```{r}
fars_summarize_years(c(2013,2014,2015))
```

## `fars_map_state` function
It displays a map of the inpute state of US for these have accidents by location and yearly.
```{r}
state.num<-20
year_input<-2013
fars_map_state(state.num,year_input)
```
